Bayesian exponential random graph modeling of KWL Burial network

### 2020-10-22 analysis/paper/paper.Rmd
edit code and method section

# Kiwulan in northeastern Taiwan

@@

pre_E <- burial_three_period_age_number$Phase[4]
post_E <- burial_three_period_age_number$Phase[3]
```

@@

The earliest record of the use of prestige goods in burials can be traced back to the time when the European arrival in the 17th century according to historical records from the Spanish [@Borao2009; @LiandWu2006]. For example, the document written by a Spanish Priest described that the Spanish soldier used carnelian beads to exchange natural resources with local Indigenous people since their high values in Indigenous culture [@LiandWu2006]. In addition, there was a custom that people would bury imported beads and ceramics with deceased to display their status and influence [@LiandWu2006]. Despite we observed the uneven distribution of prestige goods in burials, the degree of differentiation across burials over time remain unclear. The discord over the discussion of Kiwulan social organization could be associated with chronological differences that has not yet been well identified. That is, the observed uneven distribution may be explained as result of social change from less social differences to more social differences when Indigenous societies involved in the complex trade network stimulated by the Europeans. We assume that the changes would be supported by chronological differences in the structure of burial networks. Analyzing and comparing networks between different phases relative to cultural contact will give an insight into the degree and changes in social differences. This is important to understand the reactions of Kiwulan residents in the context of indirect impacts of European colonization.

<!--chronology-->
We assigned burials to the pre-European period (n = `r pre_E`), European and post-European period (n = `r post_E`) for network comparison according to depth, radiocarbon dates, diagnostic materials, stratigraphic data, and previous studies. Since the sedimentary characteristics show vertical and lateral continuity cross units of AD section, depth is used as the first criterion for burial chronology. The chronology is then evaluated by the burials with time marker, light gray glazed jars, also called An-ping jars. The jars were believed to be produced in Southeast China and circulated in Southeast Asia from the late 16th to 18th century, but are often found at sites with the European occupation in the early 17th century in Taiwan [@Berrocal2018; @Hsieh1995; @Ketel2011]. In addition to the diagnostic jars, three radiocarbon dates of charcoal associated with three different burials are used as another evidence to evaluate the chronology [@Chen2007]. Some burials that were heavily disturbed by modern construction are excluded to avoid misinterpreting due to incomplete data. With the information of burials, including quantity of grave goods, types of grave goods, burial identities, we can reconstruct social relations between individuals by linking them according to similar prestige goods. This is based on the assumption that the prestige goods imply social practices of trade, exchange, and gifting that reflect social relations between individuals [@Coward2013, pp. 252]. Also, prestige materials usually relate to a representation of social status, where the same economic or social class, tend to use similar goods as a status indicators [citations]. Therefore, we build network where burials represent actors (nodes in the network) that are linked when they have th same prestige goods in common. The prestige goods we identified including agate beads, gold-foil beads, imported Chinese porcelains, gold foils, and fish-shaped ornaments, according to previous research [@Cheng2008; @Hsieh2009; @Hsieh2012; @Wang2011] and the description in historical records [@Borao2009; @LiandWu2006].

@@

We analyzed burial data from published excavation reports, and original fieldwork notes of Kiwulan. We do not include the burials of the Chinese phase in our network analysis due to small sample size. We assume that the pericolonial impact hypothesis would be supported by changes in network structures build upon flow of resources, prestige goods here. In our case, burials are nodes that are linked when they have goods in common. Given that gold-foil beads and agate beads commonly found across burials with substantial differences in counts, we assigned those prestige goods into three levels, high, medium, and low, to indicate quantity based on the distribution of their counts. That means if burial 1 and burial 2 both have high quantity of agate beads, then there will be a tie connecting them. For less frequent prestige goods, including imported porcelains, gold foils, and fish-shaped ornaments, we assign a tie between burials when they possess each type of goods in common. Node attributes here include osteological data and mortuary practices, such as age, gender, ritual practice, and overall burial goods quantity index. Ritual practice is identified by locally made ceramics placed above graves that suggests funeral feasting according to historical records. Quantity index is an important attribute for economical inequality that we assigned the burials into three classes, high, medium, and low, as an index of wealth, based on the total number of burial goods. Since burials tend to have multiple prestige goods in common, the network ties are weighted instead of binary data (where the value 1 represents a tie and the value 0 otherwise) [@Snijders2011]. For example, if two burials have both low quantity of glass beads and porcelain in common, the tie is given a value of 2. Our networks are non-directed, which means ties have no orientation forming the relationship between actors to indicate a mutual relationship.

>>>>

# Burial data from Kiwulan, NE Taiwan

@@

pre_E <- burial_three_period_age_number$Phase[4]
post_E <- burial_three_period_age_number$Phase[3]
chi <- burial_three_period_age_number$Phase[1]
```

@@

The earliest record of the use of prestige goods in burials can be traced back to the time when the European arrival in the 17th century according to historical records from the Spanish [@Borao2009; @LiandWu2006]. For example, the document written by a Spanish Priest described that the Spanish soldier used carnelian beads to exchange natural resources with local Indigenous people since their high values in Indigenous culture [@LiandWu2006]. In addition, there was a custom that people would bury imported beads and ceramics with deceased to display their status and influence [@LiandWu2006]. Despite we observed the uneven distribution of prestige goods in burials, the degree of differentiation across burials over time remain unclear. The discord over the discussion of Kiwulan social organization could be associated with chronological differences that has not yet been well studied. That is, the observed uneven distribution may be explained as a result of social change from less social inequality to more social inequality when Indigenous societies in Northeastern Taiwan were involved in the complex trade network stimulated by the Europeans. We assume that the changes would be supported by chronological differences in the structure of burial networks. Analyzing and comparing networks between different phases relative to cultural contact will give an insight into changes in social differences. This is important to understand the reactions of Kiwulan residents in the context of indirect impacts of European colonization.

<!--chronology-->
To compare burial networks, we assigned burials to the pre-European period (n = `r pre_E`), European and post-European period (n = `r post_E`) for network comparison according to depth, radiocarbon dates, diagnostic materials, stratigraphic data, and previous studies. We excluded the burials from the Chinese phase (n = `r chi`) for analysis due to much smaller sample size. Since the sedimentary characteristics show vertical and lateral continuity cross units of AD section, depth is used as the first criterion for burial chronology. The chronology is then evaluated by the burials with light gray glazed jars, also called An-ping jars that were believed to be produced in Southeast China. They circulated in Southeast Asia from the late 16th to 18th century, and often found at sites with the European occupation in the early 17th century in many Aisan places, including Taiwan [@Berrocal2018; @Hsieh1995; @Ketel2011].

In addition to the diagnostic jars, three radiocarbon dates of charcoal associated with three different burials are used as another evidence to evaluate the chronology [@Chen2007]. Some burials that were heavily disturbed by modern construction are excluded to avoid misinterpreting due to incomplete data. With the information of burials, including quantity of grave goods, types of grave goods, burial identities, we can reconstruct social relations between individuals by linking them according to similar prestige goods. This is based on the assumption that the prestige goods imply social practices of trade, exchange, and gifting that reflect social relations between individuals [@Coward2013, pp. 252]. Also, prestige materials usually relate to a representation of social status, where the same economic or social class, tend to use similar goods as a status indicators [citations]. Therefore, we build network where burials represent actors (nodes in the network) that are linked when they have th same prestige goods in common. The prestige goods we identified including agate beads, gold-foil beads, imported Chinese porcelains, gold foils, and fish-shaped ornaments, according to previous research [@Cheng2008; @Hsieh2009; @Hsieh2012; @Wang2011] and the description in historical records [@Borao2009; @LiandWu2006].

@@

We analyzed burial data collected from the published excavation reports, and the original fieldwork notes of Kiwulan. We assume that the pericolonial impact hypothesis would be supported by changes in network structures build upon flow of resources, prestige goods here. We treat the presence of relations between burials based on number or types of  prestige items from regional trade network. where a few prestige goods controlled by a small number of individuals. This is because prestige items could represent a relationship, where a few prestige goods controlled by a small number of individuals. Based on this assumption, we treat the presence of relations between burials through the burial goods in comoon.  Given that gold-foil beads and agate beads commonly found across burials with substantial differences in counts, we classifed those beads into three levels, high, medium, and low, to indicate quantity based on the distribution of their counts. That means if burial 1 and burial 2 both have high quantity of agate beads, then there will be a tie connecting them. For less frequent prestige goods, including imported porcelains, gold foils, and fish-shaped ornaments, we linked two burials when they possess each type of goods in common. Node attributes here include osteological data and mortuary practices, such as age, gender, ritual practice, and overall burial goods quantity index. Ritual practice is identified by locally made ceramics placed above graves that suggests funeral feasting according to historical records. Quantity index is an important attribute for economical inequality that we assigned the burials into three classes, high, medium, and low, as an index of wealth, based on the total number of burial goods. Since burials tend to have multiple prestige goods in common, the network ties are weighted instead of binary data (where the value 1 represents a tie and the value 0 otherwise) [@Snijders2011]. For example, if two burials have both low quantity of glass beads and porcelain in common, the tie is given a value of 2. Our networks are non-directed, which means ties have no orientation forming the relationship between actors to indicate a mutual relationship.
